---
title: "Initial Korean covid 19"
author: "Robert Solis"
date: "6/21/23"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### This data was retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Your Question - Is dying independent of sex?

## methods
  We will analyze the sex of the Korean patients and the result of whether they are alive or deceased. Ill the variables in Cat~cat, for this problem we can use the helper file to copy the neccesary code and run the data to find the correct answer.


  

## Graphical results - use ggplot commands to do this  

### Plot barplots gg way

  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=Sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=Sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```




  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
The percentages show that more females are alive after having covid then men by 56 to 44 Percent alive. While dead is 36% females to a 64% men are dead. In the overall population is 1.4 percent of females had died and 3 percent of men were dead out of the 100%.

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```
In the group 1817 and 1411, overall the female sex survived more then men. So the p value is 1 out of a thousand, it is a small chance of a male and female dying. In fishers test 9 out of 10,000 p value is essentially  the same. This shows us the ratio of  females living over males dying.

## Conclusions

In today lesson i learned how to use helper file cat~cat to take the covid take from korea and compare which sex was most affected. By changing the data in the helper file to the data in kRdata the computer sorted and created the data chart necessary to show me the results from Fisher exact test. Thus i learned men in korea died overall 2.6 percent more then women in ther overall population.
